Reliable feature point matching without the use of geometric constraints

### 2013-06-03 text/clustering.tex
Added more text and a few images

\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\title{Clustering feature points}
\author{Jonas Arnfred}

@@

\begin{abstract}
% TODO!
\end{abstract}

@@

\section{Experimental Setup}
\label{experiment}

@@

With this in mind I have created a test suite by cropping out $300$ 
pixel by $300$ pixel patches of the original pair with a random vertical 
and horizontal offset. Given an image pair I can obtain $n$ pairs of 
patches where the two patches in any pair might or might not overlap.  
This ensures that the patch pairs where there is no overlap still retain 
a basic similarity to each other, while the patches that do overlap 
often only overlap on a small part of their surface.

@@

\bibliographystyle{abbrv}
\bibliography{bibliography}
\end{document}

>>>>

%\documentclass[12pt,journal,compsoc]{IEEEtran}
\documentclass{article}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\title{Reliable feature point matching without the use of geometric 
constraints}
\author{Jonas Arnfred}

@@

\begin{abstract}
Many algorithms have been proposed to solve the problem of matching 
feature points in two or more images using geometric assumptions to 
increase the robustness of the matching. However very little work 
addresses the case where these assumptions might not hold true. In 
particular few methods address the problem of reliable matching in cases 
where we don't know if two images have any corresponding areas or 
objects. In this paper we propose two algorithms for matching feature 
points without the use of geometric constraints. The first makes use of 
the fact that any match between two images should be better than all 
possible matches within the image itself. The second algorithm extends 
this idea by using the community structure in the similarity graph of 
feature points to reliably find correspondences. To evaluate the 
algorithms experimentally we introduce an image set of 8 image pairs all 
containing view port changes and a simple method to generate a large 
amount of test cases based on these 8 pairs. The experimental results 
show that the proposed algorithm, overall, is superior to traditional 
approaches in finding correct correspondences.
\end{abstract}

@@

\section{Experiments}
\label{experiment}

@@

With this in mind I have created a test suite by cropping out $300$ 
pixel by $300$ pixel patches of the original pair with a random vertical 
and horizontal offset. Given an image pair I can obtain $n$ pairs of 
patches where the two patches in any pair might or might not overlap.  
This ensures that the patch pairs where there is no overlap still retain 
a basic similarity to each other, while the patches that do overlap 
often only overlap on a small part of their surface. Figure 
\ref{fig:fairey} shows an example of possible pairs of test images 
(\ref{fig:fairey_crop}) produced from two source images 
(\ref{fig:fairey1} and \ref{fig:fairey2}). Producing $n$ such pairs 
allows us to test not just how well the matching algorithm performs on 
corresponding images, but also how many false positives we get on 
similar images that don't overlap.

The set of source images and homographies as well as the a script to 
generate the cropped test sets based on them have been published. The 
actual testsets used for testing in this paper are also available 
online\footnote{They can be found on INSERT URL HERE}.

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.50\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example1}
			\caption{Mural by Fairey (1)}
			\label{fig:fairey1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.50\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example2}
			\caption{Mural by Fairey (2)}
			\label{fig:fairey2}
		\end{subfigure}%
	}%
	\\
	\makebox[\textwidth][c]{
		%~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)

		\begin{subfigure}[t]{1.16\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/crop_examples}
			\caption{Image pairs}
			\label{fig:fairey_crop}
		\end{subfigure}%
	}%
	\label{fig:fairey}
	\caption{Example of test set produced by pair of images}
\end{figure}

\subsection{Experimental setup}

Using the image test sets described above, the \emph{MM} and \emph{MMC} 
algorithms have been tested against a baseline algorithm incorporating 
ideas from the algorithm Lowe proposed in \cite{lowe2004sift} as well as 
the gold standard algorithm for calculating homographies proposed in 
\cite{multipleView}. For reference the algorithm can be seen in 
\ref{alg:baseline}.

\begin{algorithm}
\caption{Baseline Match Algorithm (\emph{BM})}
\label{alg:baseline}
\begin{algorithmic}
\Require $image_1$ : image, $image_2$ : image, $threshold$ : float $\in 
\mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final}\gets \varnothing$
\State $features_1 \gets getFeatures(image_1)$
\State $features_2 \gets getFeatures(image_2)$
\ForAll{$f_i \in features_1$} \Comment Get best match for all features 
in $image_1$
	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features_2)$
	\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
	f_m)$
	\If{$ratio < threshold$}
		\State $matches_{init} \gets \left(f_i, f_m\right)$
	\EndIf
\EndFor
%
\ForAll{$f_i \in features_2$} \Comment Get best match for all features 
in $image_2$
	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features_1)$
	\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
	f_m)$
	\If{$ratio < threshold$}
		\State $matches_{init} \gets \left(f_i, f_m\right)$
	\EndIf
\EndFor
%
\ForAll{$\left(f_i, f_j \right) \in matches_{init}$} \Comment Filter 
matches
\If{$\left(f_j, f_i \right) \in matches_{init} \wedge \left(f_j, 
f_i\right) \not\in matches_{final}$}
		\State $matches_{final} \gets (f_i, f_j)$
	\EndIf
\EndFor \\
\Return $matches_{final}$
\end{algorithmic}
\end{algorithm}

For each test set of cropped image pairs the three 
algorithms\footnote{\emph{MM}, \emph{MMC} and \emph{BM}} are applied to 
each pair of images, and the threshold for each is varied over 50 evenly 
spaced values. For \emph{MM} and \emph{BM} these values lie between 
$0.4$ and $0.8$ where as the threshold for \emph{MMC} is varied between 
$0.8$ and $1.0$. For each match returned by any of the algorithms, the 
match is registered as an inlier if the endpoint in the second image 
lies within $5$px of the true position calculated by the homography.  
That is, given $H$ relating $I_1$ with $I_2$ and a match ($p_1, p_2$) 
then the match is an inlier if $\left\vert Hp_1 - p_2 \right\vert < 5$.  
This is a relatively lenient threshold since radial distortion from the 
lens as well as inconsistencies to the planar assumption of the 
background means that $H$ can not be entirely accurate. The experiments 
are run on the $8$ image pairs in the test set, as well as the graf 
image from \cite{mikolajczyk2005performance}.

@@

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/fairey_burma_1}
			\caption{Burma by Fairey (1)}
			\label{fig:burma1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/fairey_burma_2}
			\caption{Burma by Fairey (2)}
			\label{fig:burma2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_fairey_burma}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_burma}
		\end{subfigure}%
	}%
	\label{fig:burma}
	\caption{results on Burma by Fairey}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.25\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/fairey_lady_1}
			\caption{Lady by Fairey (1)}
			\label{fig:lady1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.25\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/fairey_lady_2}
			\caption{Lady by Fairey (2)}
			\label{fig:lady2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.55\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_fairey_lady}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_lady}
		\end{subfigure}%
	}%
	\label{fig:lady}
	\caption{results on Lady by Fairey}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/scharf_1}
			\caption{Faces by Scharf (1)}
			\label{fig:faces1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/scharf_2}
			\caption{Faces by Scharf (2)}
			\label{fig:faces2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_scharf}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_faces}
		\end{subfigure}%
	}%
	\label{fig:faces}
	\caption{results on Faces by Scharf}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/houston_1}
			\caption{Houston (1)}
			\label{fig:houston1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/houston_2}
			\caption{Houston (2)}
			\label{fig:houston2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_houston}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_houston}
		\end{subfigure}%
	}%
	\label{fig:houston}
	\caption{Results on Houston}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.25\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/blu_head_1}
			\caption{head by blu (1)}
			\label{fig:head1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.25\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/blu_head_2}
			\caption{head by blu (2)}
			\label{fig:head2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.55\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_blu_head}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_head}
		\end{subfigure}%
	}%
	\label{fig:head}
	\caption{results on head by blu}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.25\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/blu_pencil_1}
			\caption{Pencil by Blu (1)}
			\label{fig:pencil1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.25\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/blu_pencil_2}
			\caption{Pencil by Blu (2)}
			\label{fig:pencil2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.55\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_blu_pencil}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_pencil}
		\end{subfigure}%
	}%
	\label{fig:pencil}
	\caption{Results on Pencil by Blu}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/banksy_city_1}
			\caption{City by Banksy (1)}
			\label{fig:city1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/banksy_city_2}
			\caption{City by Banksy (2)}
			\label{fig:city2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_banksy_city}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_city}
		\end{subfigure}%
	}%
	\label{fig:city}
	\caption{results on City}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/banksy_stroller_1}
			\caption{Stroller by Banksy (1)}
			\label{fig:stroller1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/banksy_stroller_2}
			\caption{Stroller by Banksy (2)}
			\label{fig:stroller2}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_banksy_stroller}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_stroller}
		\end{subfigure}%
	}%
	\label{fig:stroller}
	\caption{Results on Stroller by Banksy}
\end{figure}

\begin{figure}
	\makebox[\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/graf_1}
			\caption{Graf (1)}
			\label{fig:graf1}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/graf_3}
			\caption{Graf (3)}
			\label{fig:graf3}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_graf}
			\caption{X: Nb of correct matches. Y: Accuracity}
			\label{fig:result_graf}
		\end{subfigure}%
	}%
	\label{fig:graf}
	\caption{results on Graf}
\end{figure}

In figure \ref{fig:stroller} to \ref{fig:graf} the results on the images

\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}
